Gelation as condensation frustrated by hydrodynamics and mechanical tension

### 2017-01-31 manuscript4.tex
Try to integrate Lambda max

\begin{figure}
\includegraphics{figs/microrheology2P.pdf}
\caption{\textbf{Time-dependent microrheology} of three percolating samples.
By increasing density from top to bottom $\phi=8,\,16,\,27~\%$, $c_p=1.5,\,1.2,\, 1$ g/$\mathrm{\ell}$. 
The thick gray curve is the internal stress $\Sigma$ obtained from the measure of bond-breaking probability. 
Elastic ($G^\prime$) and viscous ($G^{\prime\prime}$) shear moduli at high frequency ($f=0.1\tau_\mathrm{B}^{-1}$), obtained by two-particle microrheology, are drawn respectively as red filled and black open circles. 
The dotted vertical line shows the percolation time.
}
\label{fig:microrheology2P}
\end{figure}

@@

In Figure~\ref{fig:microrheology2P}, to relate the timing of stress relaxation, mechanical gelation and percolation, we superimpose the time evolutions of internal stress and both parts of the shear moduli at the highest accessible frequency ($f=0.1\tau_\mathrm{B}^{-1}$). The time needed for $\Sigma(t)$ to relax completely seems to be approximatively independent of colloid volume fraction. Moreover, stress relaxation occurs independently of percolation. In dilute gels, Figure~\ref{fig:microrheology2P}a, internal stresses are relaxed when percolation takes place, whereas in dense gels, Figure~\ref{fig:microrheology2P}c, stress relaxation occurs mostly after percolation. Percolation is obviously a necessary condition for mechanical gelation, since the crossing of $G^\prime$ and $G^{\prime\prime}$ occurs always later than percolation. However there is a long delay between percolation and the rise of $G^\prime$, therefore percolation is not enough to ensure mechanical stability. In dense gel, Figure~\ref{fig:microrheology2P}c, the just percolated network rearranges under internal stresses until a mechanically stable configuration is found. For dilute gels, Figure~\ref{fig:microrheology2P}a, we observe larger fluctuations on $\Sigma(t)$ after percolation, probably due to the release of local internal stresses progressively leading to a mechanically stable configuration.

@@

Figure~\ref{fig:breaking}a shows a strand rupture reconstructed from confocal 3D images, the colours mapping the intensity of internal stresses acting on each particle. In practice, the attraction potential well is too narrow relative to the particle localisation errors for us to extract directly the force acting on a particle. Therefore, we use the elongation of the particle neighbourhood as a proxy. The more the bonds around a particle are oriented along a single axis, the more tension is exerted on the particle. This local elongation is quantified using bond orientational order~\cite{steinhardt1983boo} for the 2-fold symmetry $q_2$, see Methods for a precise definition. In Fig.~\ref{fig:breaking}a we observe the slow build up of tension over the first 30 min, until the tension is concentrated on a weak part, a one particle thin strand. This strand breaks and the particles involved merge in their respective part of the network that then stay apart from each other. This is the only elementary process allowed for a percolated network to lower the free energy and it is a mechanically driven coarsening process \cite{Tanaka2007}. The driving force is simply attractive interactions between particles favouring compact structures, which can be rephrased by 
interfacial energy in a coarse-grained picture.

>>>>

\begin{figure}
\includegraphics{figs/microrheology2P.pdf}
\caption{\textbf{Time-dependent mechanical response} of three percolating samples.
By increasing density from top to bottom $\phi=8,\,16,\,27~\%$, $c_p=1.5,\,1.2,\, 1$ g/$\mathrm{\ell}$. 
The thick gray curve is the internal stress $\Sigma$ obtained from the measure of bond-breaking probability. 
Elastic ($G^\prime$) and viscous ($G^{\prime\prime}$) shear moduli at high frequency ($f=0.1\tau_\mathrm{B}^{-1}$), obtained by two-particle microrheology, are drawn respectively as filled and open circles. 
The dark red bars show $\Lambda_\mathrm{max}$, the size of the largest strand breaking event.
The dotted vertical line shows the percolation time.
}
\label{fig:microrheology2P}
\end{figure}

@@

In Figure~\ref{fig:microrheology2P}, to relate the timing of stress relaxation, mechanical gelation and percolation, we superimpose the time evolutions of internal stress and both parts of the shear moduli at the highest accessible frequency ($f=0.1\tau_\mathrm{B}^{-1}$). The time needed for $\Sigma(t)$ to relax completely seems to be approximatively independent of colloid volume fraction. Moreover, stress relaxation occurs independently of percolation. In dilute gels, Figure~\ref{fig:microrheology2P}a, internal stresses are relaxed when percolation takes place, whereas in dense gels, Figure~\ref{fig:microrheology2P}c, stress relaxation occurs mostly after percolation. Percolation is obviously a necessary condition for mechanical gelation, since the crossing of $G^\prime$ and $G^{\prime\prime}$ occurs always later than percolation. However there is a long delay between percolation and the rise of $G^\prime$, therefore percolation is not enough to ensure mechanical stability. In dense gel, Figure~\ref{fig:microrheology2P}c, the just percolated network rearranges under internal stresses until a mechanically stable configuration is found. %For dilute gels, Figure~\ref{fig:microrheology2P}a, we observe larger fluctuations on $\Sigma(t)$ after percolation, probably due to the release of local internal stresses progressively leading to a mechanically stable configuration.

@@

To understand how a percolated suspension evolves to a mechanically stable gel, we complement the mean field measurements above with local observables.
Figure~\ref{fig:breaking}a shows a strand rupture reconstructed from confocal 3D images, the colours mapping the internal strain acting on each particle. 
In practice, the attraction potential well is too narrow relative to the particle localisation errors for us to measure reliably the strain of a single bond.
Therefore, we measure the elongation of the particle neighbourhood. 
%The more the bonds around a particle are oriented along a single axis, the more tension is exerted on the particle. 
This local elongation is quantified using bond orientational order~\cite{steinhardt1983boo} for the 2-fold symmetry $q_2$, see Methods for a precise definition. In Fig.~\ref{fig:breaking}a we observe the slow build up of tension over the first 30 min, until the tension is concentrated on a weak part, a one particle thin strand. This strand breaks and the particles involved merge in their respective part of the network that then stay apart from each other. This is the only elementary process allowed for a percolated network to lower the free energy and it is a mechanically driven coarsening process~\cite{Tanaka2007}. The driving force is simply attractive interactions between particles favouring compact structures, which can be rephrased by 
interfacial energy in a coarse-grained picture.
